A simplified dynamic bioenergetic model for coral-\emph{Symbiodinium} symbioses

### 2016-10-06 JTB_ms/JTB_ms.tex
updates from UCSB

The nutritional exchange between corals and \emph{Symbiodinium} directly
underlies the capacity of corals to build coral reef ecosystems, worth
trillions of US Dollars annually (Costanza, Groot, and Sutton 2014).
However, the complex symbiotic metabolism of corals is vulnerable to
disruption by numerous anthropogenic environmental perturbations,
jeopardizing their future persistence. In order to understand and
predict coral responses to complex changes in the environment, a
mechanistic understanding of how multiple interacting factors drive the
individual and emergent physiology of both symbiotic partners is
necessary. Such a task is well suited for theoretical modeling
frameworks such as Dynamic Energy Budget (DEB) theory (Kooijman 2010),
although the complexity of such theory makes these efforts inaccessible
to many biologists (Jager, Martin, and Zimmer 2013). In order to bridge
this gap, we present here a simplified dynamic bioenergetic model for
coral-\emph{Symbiodinium} symbioses that aims to mechanistically
integrate the impacts of complex environmental change on the
physiological ecology of reef corals.

In reef coral symbioses, intracellular \emph{Symbiodinium} translocate
photosynthetically-fixed carbon to support coral metabolism, and utilize
the animal's metabolic waste products, including nitrogenous compounds
and carbon dioxide, in return (Muscatine and Porter 1977). Previous
application of DEB theory to this syntrophic system (Muller et al. 2009)
demonstrated a stable symbiotic relationship and qualitatively realistic
growth and biomass ratios across gradients of ambient irradiance,
nutrients, and food. This model assumed that 1) \emph{Symbiodinium} has
priority access to carbon through photosynthesis, 2) the coral animal
has priority access to dissolved nitrogen through contact with seawater,
and 3) each partner shares with the other only what it cannot use
itself. In its simplest form, this principle of sharing the surplus
sufficiently describes diverse syntrophic interactions among organs and
organisms (e.g., trees, duckweeds, corals), suggesting the mechanism is
mathematically and evolutionarily robust (Nisbet et al., submitted).

While the work of Muller et al. (2009) applying DEB formalism to this
system represents the most significant theoretical contribution in coral
symbiosis research to date, we aim to strengthen the role of theory and
broaden its potential application to corals in three primary ways:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \emph{Develop a detailed module of environmental stress.} Of primary
  interest to coral biologists and ecologists is symbiosis dysfunction
  under environmental stress, resulting in coral ``bleaching''--the loss
  of algal symbionts from the association (Jokiel and Coles 1977).
  Photooxidative stress in \emph{Symbiodinium} is considered a primary
  trigger of bleaching in response to high temperature and/or light
  (Weis 2008), and prolonged or severe bleaching can result in
  mortality, though corals sometimes recover their symbionts. Bleaching
  susceptibility, severity, and recovery may by influenced by
  interacting factors such as heterotrophy and nutrient availability
  (Wooldridge 2014b), and the genetic identity of \emph{Symbiodinium}
  (Glynn et al. 2001). To simulate these bleaching-related phenomena, we
  develop a generalized framework linking overreduction of the
  photosynthetic light reactions to downstream impacts of
  photoinhibition and photodamage.
\item
  \emph{Reduce theoretical and mathematical complexity.} Following the
  logic of Jager, Martin, and Zimmer (2013), we exclude certain features
  of formal DEB models in order to capture behaviors of interest with
  the simplest possible formulation. Here, we present a model without
  reserves, maturity, or reproduction (see Kooijman 2010). This
  formulation restricts the model's scope to the bioenergetics of growth
  and symbiosis dynamics in adult corals, but greatly reduces
  theoretical complexity and parameter numbers, which is advantageous
  given the relative paucity of data for corals. However, our primary
  motivation for reducing complexity was to increase accessibility and
  applicability for biologists and ecologists without requiring
  significant expertise in DEB theory.
\item
  \emph{Provide well-documented, open-access code.} In order to
  facilitate the continued development and application of theoretical
  modeling tools for coral symbioses, we provide open access to the
  model in the form of detailed, commented code written in the R
  language (R Core Team 2014). With an accessible and modular framework,
  we envision this as a resource for futher development by the
  scientific community to include additional complexity and
  problem-specific components. The R language was chosen because it is
  freely available and in common use by biologists and ecologists, to
  widen the audience for this work.
\end{enumerate}

@@

In this model of coral-algal symbiosis, carbon and nitrogen are acquired
by each partner and used to construct biomass. A graphical
representation of the model is presented in Fig. 1. We use C-moles as
the unit of biomass for consistency with the rigorous mass balance of
DEB theory: 1 C-mole is equivalent to the amount of biomass containing 1
mole of Carbon atoms. Host biomass (\(H\)), symbiont biomass (\(S\)),
and prey biomass (\(X\)) have fixed, but different, molar N:C ratios
(Table 1). Carbon and nitrogen are combined to produce biomass by
synthesizing units (SU), which are mathematical specifications of the
formation of a product from two substrates; we use the ``parallel
complementary'' formulation of Kooijman (2010) to specify these fluxes.
The two state variables in this dynamical system are symbiont biomass
and coral biomass; because resources are acquired proportionally to
surface area (and surface area is assumed proportional to volume for
corals (i.e., they are ``V1-morphs'' in DEB terminology (Kooijman
2010))), biomass increases exponentially during growth. The rate of
increase in coral biomass (i.e.~growth) and the ratio of symbiont to
host biomass are the (i.e.~symbiosis dynamics) are the responses of
interest of the system. Below we describe the formulation of each flux
involved in producing these responses.

Table 1

@@

d\textsuperscript{-1}

@@

d\textsuperscript{-1}

@@

@@

Non-photochemical quenching capacity of symbiont

@@

40

@@

Excess photon energy that doubles ROS prod.

@@

d\textsuperscript{-1}

@@

Also include tables of fluxes and environmental inputs?

\begin{figure}[htbp]
\centering
\includegraphics{../img/model.png}
\caption{Graphical representation of coral-algal symbiosis model.}
\end{figure}

@@

The coral animal acquires both carbon and nitrogen from feeding on prey
from the environment. Prey acquisition is specified by Michaelis-Menten
kinetics using a maximum area-specific feeding rate and half-saturation
constant:

@@

Additionally, the coral animal can acquire nitrogen from the surrounding
seawater. This nitrogen source is assumed to represent ammonium, the
primary form utilized by corals (Wang and Douglas 1998; Yellowlees,
Rees, and Leggat 2008). The uptake of nitrogen from the environment is
specified by Michaelis-Menten kinetics using a maximum area-specific
uptake rate and half-saturation constant:

@@

Coral biomass formation is specified by a parallel complementary SU that
combines carbon and nitrogen to form biomass. In addition to the carbon
and nitrogen acquired by direct uptake (Eqs. 1 and 2), the coral
recycles a portion of the nitrogen liberated by biomass turnover
(\(r_{NH}=\sigma_{NH}j_{HT}^0\)), and receives surplus fixed carbon
shared by the symbiont (\(\rho_C\)), such that the total biomass
formation is specified as:

@@

The amount of nitrogen input to the coral biomass SU in excess of what
is actually consumed in biomass formation (i.e., surplus nitrogen, or
the ``rejection flux'' in SU terminology) is then made available to the
symbiont, and is specified as:

@@

Due to the inherent inefficiency of the parallel complementary SU
formulation, there will always be some nitrogen shared with the symbiont
even when coral biomass formation is strongly nitrogen-limited.
Likewise, there is always a non-zero rejection flux of carbon from the
coral biomass SU, which is assumed to be lost to the environment.

@@

The symbiont produces fixed carbon through photosynthesis, a process
represented here by a single SU with two substrates: light (photons) and
inorganic carbon (CO\textsubscript{2}). The amount of light absorbed by
the symbiont depends on the scalar irradiance at the site of light
absorption, which is modified substantially relative to external
downwelling irradiance owing to multiple scattering by the coral
skeleton and self-shading by surrounding symbionts (Enríquez, Méndez,
and Iglesias-Prieto 2005; Marcelino et al. 2013). We used data from
Marcelino et al. (2013) to empirically derive the ratio of internal
scalar irradiance to external downwelling irradiance as a function of
symbiont density (expressed as symbiont to host biomass ratio), and
subsequently multiply this quantity by the external downwelling
irradiance \(L\) and the effective light-absorbing surface area of
symbiont biomass \(ds\) to specify the amount of light absorbed:

\begin{equation} j_L = \big[1.26 + 1.39 \cdot \exp(-6.48 \cdot {S \over H})\big] \cdot L \cdot \bar{a}^* \end{equation}

We then specify two pathways for input of inorganic carbon to the
photosynthesis SU: 1) passive diffusion of CO\textsubscript{2} from the
external environment, and 2) active delivery of CO\textsubscript{2} to
the symbiont by the host. The passive flux ensures that some
CO\textsubscript{2} is always available to photosynthesis, and the
active flux encompasses the potentially diverse mechanisms by which the
host may enhance CO\textsubscript{2} availability for the symbiont,
including active transport of bicarbonate, carbonic anhydrase-catalyzed
conversion of bicarbonate to CO\textsubscript{2} to promote diffusion
toward the symbiont, and acidification of the symbiosome to increase
localized CO\textsubscript{2} concentrations. Since the host physically
separates the symbiont from the external environment, both the passive
and active flux rates are proportional to host surface area. A more
rigorous, mechanistic model of inorganic carbon processing would require
spatially explicit internal pools accounting for pH and carbon
speciation, which is beyond the current scope of this work. Instead, the
specification of CO\textsubscript{2} delivery rates offers the user the
opportunity to compare different rates of CO\textsubscript{2} delivery
that may characterize different coral species (Wooldridge 2014a). The
input of CO\textsubscript{2} to the photosynthesis SU is therefore
specified as:

@@

In addition to the inputs CO\textsubscript{2} specified in Eqs. 7,
additional CO\textsubscript{2} representing the metabolic production of
CO\textsubscript{2} from host and symbiont biomass turnover
(\(r_{CH}=\sigma_{CH}j_{HT}^0; r_{CS}=\sigma_{CS}j_{ST}^0\)) is made
available to the photosynthesis SU. fixed carbon is produced by the
photosynthesis SU according to:

@@

where \(j_{CPm}\) is the maximum specific rate of photosynthesis, and
\(c_{ROS}\) is the photooxidative stress multiplier (see below).
Dividing by \(c_{ROS}\) causes the rate of photosynthesis to decline in
response to photo-oxidative stress, a phenomenon known as
photoinhibition.

@@

This excess light energy must be quenched by alternative pathways in
order to prevent photo-oxidative damage. \emph{Symbiodinium} may utilize
a variety of pathways for non-photochemical quenching (NPQ; Roth 2014),
which we collect in a total capacity for NPQ as a parameter of the
symbiont (\(j_{NPQ}\)). If light energy exceeds the total capacity of
both carbon fixation and NPQ, then damaging reactive oxygen species
(ROS) may be produced. We represent this as a scaled flux of ROS
\(c_{ROS}\), which takes a value of 1 when all light absorbed is
quenched by photochemistry and NPQ, and increases as the amount of
excess excitation energy increases.

\begin{equation} c_{ROS} = 1 + \bigg[\bigg({{j_{eL} - j_{NPQ}} \over k_{ROS}}\bigg)^k\bigg]_+ \end{equation}

where \(j_{NPQ}\), \(k_{ROS}\), and \(k\) are parameters of the symbiont
that determine the onset and rate of ROS production. Importantly,
\(c_{ROS}\) as specified here is not a function of absolute light
absorption, but rather the amount of excess light energy \(j_{eL}\)
after accounting for carbon fixation and NPQ. A direct consequence of
this formulation is that carbon-limitation of photosynthesis can lead to
photo-oxidative stress, a mechanism of biological importance (Wooldridge
2009) that was not captured by previous representations of
photo-oxidative stress (Eynaud, Nisbet, and Muller 2011). Moreover, this
formulation allows functional diversity among symbiont types to be
explored by changing the parameters \(j_{NPQ}\), \(k_{ROS}\), and \(k\).

Carbon fixed by photosynthesis (\(j_{CP}\)) is then used in conjunction
with nitrogen shared by the host (\(\rho_N\)) and a proportion of
nitrogen recycled from symbiont biomass turnover
(\(r_{NS}=\sigma_{NS}j_{ST}^0\)) to build new symbiont biomass,
following the SU equation:

@@

The rejection flux of carbon from this SU represents the amount of fixed
carbon produced by photosynthesis in excess of what can be used to
produce symbiont biomass; this surplus \(\rho_C\) is translocated to the
coral host:

@@

\begin{equation} j_{ST} = j_{ST}^0(1 + 5 \cdot (c_{ROS}-1)) \end{equation}

This second component of symbiont biomass loss can represent both
photodamage and/or symbiont expulsion (i.e., bleaching), both of which
happen in response to higher levels or ROS production. The constant 5 is
included to increase biomass loss in response to ROS. (Note that
recycling of symbiont biomass turnover (\(rNS\) and \(rCS\)) only occurs
based on the basal maintenance related turnover (i.e., \(j_{ST}^0\)),
and not the bleaching-related biomass loss, as this loss represents
biomass being damaged or being expelled from the holobiont).

\subsection{Model state equations}\label{model-state-equations}

Finally, the balance equations representing the specific growth rates of
symbiont and host biomass over time can be expressed as:

@@

A time-stepping Euler method was used to solve the state equations since
the production and rejection fluxes of the SUs are imiplicitly defined.
Specifically, the rejection fluxes of carbon and nitrogen from the
symbiont and host biomass SUs act as reciprocal input fluxes to the
other SU. In addition, the rejection flux of excitation energy from the
photosynthesis SU acts to reduce its own production flux (i.e.,
photoinhibition), and hence a discretized time-stepping procedure was
necessary. A vector of time values was created for each simulation run,
along which dynamic environmental forcing functions (irradiance, DIN,
and prey abundance) can be designed. These vectors, along with initial
values of symbiont and host biomass, then serve as input to the
time-stepping function, which solves for the current system state using
values of the previous system state where necessary. A default time step
of 0.1 days was used for all simulations, which were performed using R
code that is available in the data repository accompanying this article:
\url{github.com/jrcunning/Rcoral}.

To aid in visualizing model results, we calculated values to indicate
the degree to which product formation at an SU was limited by
availability of either of its two substrates

@@

\section{Model behavior evaluation}\label{model-behavior-evaluation}

To analyze qualitative behavior, we ran the model to steady state across
gradients of external irradiance and nutrients (Fig. 2), which revealed
patterns that are consistent with observed phenomena in corals.

>>>>

The nutritional exchange between corals and \emph{Symbiodinium} directly
underlies the capacity of corals to build coral reef ecosystems, worth
trillions of US Dollars annually (Costanza, Groot, and Sutton 2014).
However, the complex symbiotic metabolism of corals is vulnerable to
disruption by numerous anthropogenic environmental perturbations,
jeopardizing their future persistence. In order to understand and
predict coral responses to complex changes in their environment, a
mechanistic understanding of how multiple interacting factors drive the
individual and emergent physiology of both symbiotic partners is
necessary. Such a task is well suited for theoretical modeling
frameworks such as Dynamic Energy Budget (DEB) theory (Kooijman 2010),
although the complexity of such theory makes these efforts inaccessible
to many biologists (Jager, Martin, and Zimmer 2013). In order to bridge
this gap, we present here a simplified dynamic bioenergetic model for
coral-\emph{Symbiodinium} symbioses that aims to mechanistically
integrate the impacts of complex environmental change on the
physiological performance of reef corals, including responses to
environmental stress.

In reef coral symbioses, intracellular \emph{Symbiodinium} translocate
photosynthetically fixed carbon to support coral metabolism, while the
animal host provides access to inorganic nutrients and carbon dioxide
(Muscatine and Porter 1977). Previous application of DEB theory to this
syntrophic system (Muller et al. 2009) demonstrated a stable symbiotic
relationship and qualitatively realistic growth and biomass ratios
across gradients of ambient irradiance, nutrients, and food. This model
assumed that 1) \emph{Symbiodinium} has priority access to fixed carbon
through photosynthesis, 2) the coral animal has priority access to
dissolved nitrogen through contact with seawater, and 3) each partner
shares with the other only what it cannot use for its own growth. In its
simplest form, this principle of sharing the surplus is sufficient to
describe the dynamics of diverse syntrophic organs and organisms (e.g.,
trees, duckweeds, corals), suggesting the mechanism is mathematically
and evolutionarily robust (Nisbet et al., in prep.).

While the formal DEB model of Muller et al. (2009) represents the most
significant theoretical contribution in coral symbiosis research to
date, we aim to strengthen the role of theory and broaden its potential
application in three primary ways:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \emph{Develop a detailed module of environmental stress.} Of primary
  interest to coral biologists and ecologists is symbiosis dysfunction
  under environmental stress, resulting in coral ``bleaching''--the loss
  of algal symbionts from the association (Jokiel and Coles 1977).
  Photooxidative stress in \emph{Symbiodinium} is considered a primary
  trigger of bleaching in response to high temperature and/or light
  (Weis 2008), and prolonged or severe bleaching can result in
  mortality, though corals sometimes recover their symbionts. Bleaching
  susceptibility, severity, and recovery may by influenced by
  interacting factors such as heterotrophy and nutrient availability
  (Wooldridge 2014b), and the genetic identity of \emph{Symbiodinium}
  (Glynn et al. 2001). To simulate these bleaching-related phenomena, we
  develop a generalized framework linking overreduction of the
  photosynthetic light reactions to downstream impacts of
  photoinhibition and photodamage.
\item
  \emph{Reduce theoretical and mathematical complexity.} Following the
  logic of Jager, Martin, and Zimmer (2013), we exclude certain features
  of formal DEB models in order to capture behaviors of interest with
  the simplest possible formulation. Here, we present a model without
  reserves, maturity, or reproduction (see Kooijman 2010). This
  formulation restricts the model's scope to the bioenergetics of growth
  and symbiosis dynamics in adult corals (i.e., reproduction, larval
  stages, and metamorphosis are not considered), but greatly reduces
  theoretical complexity and parameter numbers, which is advantageous
  given the relative paucity of data for corals. However, our primary
  motivation for reducing complexity was to increase accessibility and
  applicability for biologists and ecologists without requiring
  significant expertise in DEB theory.
\item
  \emph{Provide well-documented, open-access code.} In order to
  facilitate the continued development and application of theoretical
  modeling tools for coral symbioses, we provide open access to the
  model in the form of detailed, commented code written in the R
  language (R Core Team 2014). With an accessible and modular framework,
  we envision this as a resource for futher development by the
  scientific community to include additional complexity and
  problem-specific components. To widen the audience for this work, we
  chose the R language because it is freely available and in common use
  by biologists and ecologists.
\end{enumerate}

@@

In this dynamic bioenergetic model of coral-algal symbiosis, carbon and
nitrogen are acquired by each partner and used to construct biomass. A
graphical representation of the model is presented in Fig. 1, and each
model flux and parameter are defined in Tables 1 and 2, respectively. We
use C-moles as the unit of biomass for consistency with the rigorous
mass balance of DEB theory: 1 C-mole is equivalent to the amount of
biomass containing 1 mole of Carbon atoms. Host biomass (\(H\)),
symbiont biomass (\(S\)), and prey biomass (\(X\)) have fixed, but
different, molar N:C ratios (Table 2). Carbon and nitrogen are combined
to produce biomass by synthesizing units (SU), which are mathematical
specifications of the formation of a product from two substrates; we use
the parallel complementary formulation of Kooijman (2010) to specify
these fluxes. The two state variables are symbiont biomass and coral
biomass; because resources are acquired proportionally to surface area,
and surface area is proportional to volume (i.e., corals are
``V1-morphs'' in DEB terminology (Kooijman 2010)), biomass increases
exponentially during growth (indeed, corals grow exponentially (Bak
1976)). The specific growth rate and the ratio of symbiont to host
biomass are the responses of interest of the system. Below we describe
the formulation of each flux involved in producing these responses.

Table 1. Model fluxes
\begin{minipage}[b]{0.12\columnwidth}\raggedright\strut
\begin{minipage}[b]{0.26\columnwidth}\raggedright\strut
Eq. no.
\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
\(j_X\)
Prey uptake rate
\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
3
\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
\(j_N\)
Nitrogen uptake rate
\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
4
\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
\(j_{HG}\)
Host biomass formation rate
\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
CmolH CmolH\textsuperscript{-1} d\textsuperscript{-1}
5
\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
\(r_{NH}\)
Recycled nitrogen from host turnover
\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
6
\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
\(\rho_N\)
Nitrogen shared with the symbiont
\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
7
\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
\(j_L\)
Light absorption rate
\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
molph CmolS\textsuperscript{-1} d\textsuperscript{-1}
9
\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
\(j_{CO_2}\)
CO\textsubscript{2} input to photosynthesis
\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
molCO\textsubscript{2} CmolH\textsuperscript{-1} d\textsuperscript{-1}
10
\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
\(r_{CH}\)
Recycled CO\textsubscript{2} from host turnover
\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
molCO\textsubscript{2} CmolH\textsuperscript{-1} d\textsuperscript{-1}
11
\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
\(r_{CS}\)
Recycled CO\textsubscript{2} from symbiont turnover
\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
molCO\textsubscript{2} CmolS\textsuperscript{-1} d\textsuperscript{-1}
12
\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
\(j_{CP}\)
Photosynthesis rate
\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
13
\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
\(j_{eL}\)
Light energy in excess of photochemistry
\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
molph CmolS\textsuperscript{-1} d\textsuperscript{-1}
14
\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
Total capacity of NPQ
\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
molph CmolS\textsuperscript{-1} d\textsuperscript{-1}
15
\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
\(c_{ROS}\)
ROS production proportional to baseline
\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
16
\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
\(r_{NS}\)
Recycled nitrogen from symbiont turnover
\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
molN CmolS\textsuperscript{-1} d\textsuperscript{-1}
17
\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
\(j_{SG}\)
Symbiont biomass formation rate
\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
CmolS CmolS\textsuperscript{-1} d\textsuperscript{-1}
18
\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
\(\rho_C\)
Fixed carbon shared with host
\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
19
\begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
\(j_{ST}\)
Symbiont turnover rate
\begin{minipage}[t]{0.26\columnwidth}\raggedright\strut
CmolS CmolS\textsuperscript{-1} d\textsuperscript{-1}
20

Table 2. Model parameters

@@

CmolH CmolH\textsuperscript{-1} d\textsuperscript{-1}

@@

CmolS CmolS\textsuperscript{-1} d\textsuperscript{-1}

@@

\(k_{NPQ}\)

@@

NPQ capacity of symbiont, relative to C-fixation

@@

4

@@

Excess photon energy that doubles ROS production, relative to baseline
levels

@@

CmolS CmolS\textsuperscript{-1} d\textsuperscript{-1}
\(b\)
Scaling parameter for strength of bleaching response
3

@@

\begin{figure}[htbp]
\centering
\includegraphics{../img/Fig1.png}
\caption{Graphical representation of coral-algal symbiosis model. Light,
CO\textsubscript{2}, prey, and DIN are acquired from the external
environment proportional to the biomass of the partner indicated by the
black box for uptake. Mass fluxes (see Table 1 for definitions) are
represented by \(j\)'s with subscripts indicating the type of mass, and
in some cases the process (e.g., \(j_{CP}\) is the flux of carbon
produced by photosynthesis), and \(\rho\)'s indicate fluxes that are
shared by one partner with the other. Parallel complementary
synthesizing units (SUs) are represented by large circles, and rejection
fluxes from these SUs are indicated by dashed lines. \(c_{ROS}\) is a
proportional rate that impacts other model fluxes by inhibition or
acceleration. Recycled mass fluxes from biomass turnover are not shown
for clarity (but see Table 1 for definitions).}
\end{figure}

\subsection{State equations}\label{state-equations}

The balance equations for symbiont and host biomass are expressed as:

These are expressed per unit of symbiont and host biomass, respectively.
The specific biomass growth and turnover rates that define these balance
equations are produced by combinations of the individual model fluxes
(see Table 1 for definitions and units), which are each expressed as
mass-specific rates (e.g., per C-mole of symbiont or host biomass). When
necessary, conversions between symbiont-mass-specific and
host-mass-specific rates are accomplished by multiplying or dividing by
the symbiont:host biomass ratio.

@@

The coral animal acquires both carbon and nitrogen from feeding on prey
from the environment. Prey acquisition is specified by Michaelis-Menten
kinetics (i.e., a Holling type II function) with a maximum feeding rate
\(j_{Xm}\) and half-saturation constant \(K_X\):

@@

Additionally, the coral animal acquires nitrogen dissolved in the
surrounding seawater, which is assumed to represent ammonium, the
primary form utilized by corals (Yellowlees, Rees, and Leggat 2008).
This gives the host (rather than the symbiont) priority in nitrogen
utilization; this capacity is supported by experimental evidence (Wang
and Douglas 1998) and is consistent with the physical arrangment of the
partners, where the host is in direct contact with the external
environment. The uptake of nitrogen from the environment is thus
specified by Michaelis-Menten kinetics using a maximum uptake rate
\(j_{Nm}\) and half-saturation constant \(K_N\):

@@

Coral biomass formation is then specified by a parallel complementary SU
that combines carbon and nitrogen, according to:

@@

where \(\rho_C\) is the surplus fixed carbon shared by the symbiont (see
below), and \(r_{NH}\) is the recycled portion of the nitrogen liberated
by host biomass turnover:

\begin{equation} r_{NH}=\sigma_{NH}j_{HT}^0 \end{equation}

The amount of nitrogen input to the coral biomass SU in excess of what
is actually consumed in biomass formation (i.e., surplus nitrogen, or
the rejection flux\footnote{Rejection fluxes must always be positive,
  and hence are specified with the notation \((x)_+\), which means
  \(max(x, 0)\).} of the SU) is then made available to the symbiont:

@@

Due to the inherent inefficiency of the parallel complementary SU
formulation, there is always some nitrogen shared with the symbiont even
when coral biomass formation is strongly nitrogen-limited. Likewise,
there is always a non-zero rejection flux of carbon from the coral
biomass SU, which is assumed to be lost to the environment.

@@

The symbiont produces fixed carbon through photosynthesis, a process
represented here by a single SU with two substrates: light (photons) and
inorganic carbon (CO\textsubscript{2}). The amount of light absorbed by
the symbiont depends on the scalar irradiance at the site of light
absorption, which is modified substantially relative to external
downwelling irradiance owing to multiple scattering by the coral
skeleton and self-shading by surrounding symbionts (Enríquez, Méndez,
and Iglesias-Prieto 2005; Marcelino et al. 2013). We used data from
Marcelino et al. (2013) to empirically derive an ``amplification
factor'' \(A\) (the ratio of internal scalar irradiance to external
downwelling irradiance) as a function of symbiont density (expressed as
symbiont to host biomass ratio), which is specified as:

\begin{equation} A = 1.26 + 1.39 \cdot \exp(-6.48 \cdot {S \over H}) \end{equation}

This amplification factor is then multiplied by the external downwelling
irradiance \(L\) and the effective light-absorbing surface area of
symbiont biomass \(\bar{a}^*\) to specify the total amount of light
absorbed:

\begin{equation} j_L =  A \cdot L \cdot \bar{a}^* \end{equation}

We then specify two pathways for input of inorganic carbon to the
photosynthesis SU: 1) passive diffusion of CO\textsubscript{2} from the
external environment (\(j_{CO_2}^p\)), and 2) active delivery of
CO\textsubscript{2} to the symbiont by the host (\(j_{CO_2}^a\)). The
passive flux ensures that some CO\textsubscript{2} is always available
to photosynthesis, and the active flux encompasses potentially diverse
mechanisms by which the host may enhance CO\textsubscript{2}
availability for the symbiont, including active transport of
bicarbonate, carbonic anhydrase-catalyzed conversion of bicarbonate to
CO\textsubscript{2} to promote diffusion toward the symbiont (Tansik,
Fitt, and Hopkinson 2015), and acidification of the symbiosome to
increase localized CO\textsubscript{2} concentrations around the
symbiont (Barott et al. 2014). Since the host physically separates the
symbiont from the external environment, both the passive and active flux
rates are proportional to host biomass. A more rigorous, mechanistic
model of inorganic carbon processing would require spatially explicit
internal pools accounting for pH and carbon speciation, which is beyond
the current scope of this work. Instead, the specification of fixed
passive and active CO\textsubscript{2} input rates, \(j_{CO_2}^p\) and
\(j_{CO_2}^a\), offers the user the opportunity to compare different
rates of CO\textsubscript{2} delivery that may characterize different
coral species (Wooldridge 2014a). The input of CO\textsubscript{2} to
the photosynthesis SU is therefore specified as:

@@

Additional metabolic CO\textsubscript{2} recycled from both the host and
symbiont biomass turnover is also made available to the photosynthesis
SU, according to:

\begin{equation} r_{CH}=\sigma_{CH}j_{HT}^0 \end{equation}\begin{equation} r_{CS}=\sigma_{CS}j_{ST}^0 \end{equation}

Fixed carbon is then produced by the photosynthesis SU according to:

@@

where \(j_{CPm}\) is the maximum specific rate of photosynthesis, and
\(c_{ROS}\) is the relative rate of reactive oxygen species production
(see below). Dividing the photosynthetic rate by \(c_{ROS}\) causes a
decline in response to photooxidative stress (i.e., photoinhibition).

@@

This excess light energy must be quenched by alternative pathways in
order to prevent photooxidative damage (Powles 1984).
\emph{Symbiodinium} utilize a variety of pathways for non-photochemical
quenching (NPQ; Roth 2014), which we collect in a relative NPQ capacity
specified as a parameter of the symbiont (\(k_{NPQ}\)). NPQ capacity may
also be susceptible to reduction by photoinhibition, so the total NPQ
capacity is specified as:

\begin{equation} j_{NPQ}= k_{NPQ} \cdot c_{ROS}^{-1} \end{equation}

If light energy further exceeds the total capacity of both carbon
fixation and NPQ, then reactive oxygen species (ROS) are produced. We
represent this as a relative quantity \(c_{ROS}\), which takes a value
of 1 when all light absorbed can be quenched by photochemistry and NPQ,
and increases as the amount of excess excitation energy further
increases:

\begin{equation} c_{ROS} = 1 + \bigg({{(j_{eL} - j_{NPQ})_+} \over k_{ROS}}\bigg)^k \end{equation}

where \(k_{NPQ}\), \(k_{ROS}\), and \(k\) are parameters of the symbiont
that determine the onset and rate of ROS production. Importantly,
\(c_{ROS}\) as specified here is not a function of absolute light
absorption, but rather the amount of excess light energy, \(j_{eL}\),
after accounting for carbon fixation and NPQ. A direct consequence of
this formulation is that CO\textsubscript{2}-limitation of
photosynthesis can lead to ROS production, an important mechanism
(Wooldridge 2009) that was not captured by previous representations of
photooxidative stress (Eynaud, Nisbet, and Muller 2011). Moreover, this
formulation allows functional diversity among symbiont types to be
explored by changing the parameters \(k_{NPQ}\), \(k_{ROS}\), and \(k\).

Carbon fixed by photosynthesis (\(j_{CP}\)) is then combined with
nitrogen shared by the host (\(\rho_N\)) and nitrogen recycled from
symbiont biomass turnover

\begin{equation} r_{NS}=\sigma_{NS}j_{ST}^0 \end{equation}

to build new symbiont biomass, following the SU equation:

@@

The rejection flux of carbon from this SU represents the amount of fixed
carbon produced by photosynthesis in excess of what can be used to
produce symbiont biomass; this surplus, \(\rho_C\), is translocated to
the coral host:

@@

\begin{equation} j_{ST} = j_{ST}^0(1 + b(c_{ROS}-1)) \end{equation}

This second component of symbiont biomass loss can represent both
photodamage and/or symbiont expulsion (i.e., bleaching), both of which
occur in response to high levels or ROS production. The parameter \(b\)
is included to scale biomass loss due to bleaching in response to ROS.
(Note that recycling of symbiont biomass turnover (\(rNS\) and \(rCS\))
only occurs based on the maintenance component of turnover (i.e.,
\(j_{ST}^0\)), and not the bleaching-related component, as this loss
represents biomass that is damaged or expelled from the host).

@@

A time-stepping Euler method was used to solve the state equations since
the production and rejection fluxes of the SUs are imiplicitly defined.
Specifically, the rejection fluxes of carbon and nitrogen from the
symbiont and host biomass SUs act as reciprocal input fluxes to the
other SU. In addition, the rejection flux of excitation energy from the
photosynthesis SU acts to reduce its own production flux (i.e.,
photoinhibition), and hence a discretized time-stepping procedure was
necessary. A vector of time values was created for each simulation run,
along which dynamic environmental forcing functions (irradiance, DIN,
and prey abundance) can be designed. These vectors, along with initial
values of symbiont and host biomass, then serve as input to the
time-stepping function, which solves for the current system state using
values of the previous system state where necessary. A time step of 0.1
days was used for all simulations, which were performed using R code
that is available in the data repository accompanying this article:
\url{github.com/jrcunning/Rcoral}.

To aid in visualizing model results, we calculated values to indicate
the degree to which product formation at an SU was limited by
availability of either of its two substrates:

@@

\section{Sensitivity analysis}\label{sensitivity-analysis}

Each default parameter value used in the model (Table 2) is based on
carefully selected relevant literature. A supplementary document
detailing the derivations of each of these default parameter values is
provided. Nevertheless, it is important to evaluate the sensitivity of
the model to changes in these parameter values. We measured fractional
change in steady state values in response to fractional change in
parameter values, relative to their default values (Table 2).
